### 2023-06-30 paper/lwfs-spl.tex
corrections_applied

For a direct spatial band-limitation,
the sound field is expanded in terms of cylindrical harmonics with respect to the target position.
The modal bandwidth was limited by truncating the expansion to a finite order.
Based on the analytical expressions of the modal impulse responses,
the LWFS driving functions are implemented as FIR filters in the time domain.
The length of the FIR filters depends on the distance between the secondary source and the target position,
ranging between 273 and 787 taps for the considered scenarios
(excluding the WFS prefilter).
%
Since this approach is based on time-domain sampling,
the resulting filters exhibit temporal aliasing artifacts,
degrading the accuracy in the frequency domain.
It is, therefore, crucial to attenuate the spectral components beyond the Nyquist limit
before the modal impulse responses are sampled.
An analytical anti-aliasing filtering is performed in the continuous-time domain
by using the 15th-order Lagrange interpolation kernel.
%
Interested readers are referred to \cite{hahn2022cylindrical} for further detail.